---
title: "FlightDelays_Naive_Bayes"
author: "Veera Namana"
date: "9/1/2019"
output: html_document
---

```{r}

flight <- read.csv("FlightDelays.csv", header = TRUE)

View(flight)

```


```{r}

str(flight)

summary(flight)

table(flight$DEP_TIME)

```

#Data Preprocessing

```{r}

sum(is.na(flight))

```

#Convert the attributes to appropriate types


```{r}
flight$Weather<- as.factor(flight$Weather)
flight$DAY_WEEK <- as.factor(flight$DAY_WEEK)
flight$Flight.Status<- as.factor(flight$Flight.Status)

str(flight)
```


#Categorize the departure time into bins based on client's needs

```{r}

flight$levels <- ifelse(flight$DEP_TIME >= 600 & flight$DEP_TIME <= 1200, "level1", ifelse(flight$DEP_TIME>1200 & flight$DEP_TIME <= 1800, "level2", ifelse(flight$DEP_TIME>1800 & flight$DEP_TIME<=2100,"level3", "level4")))

flight$levels <- as.factor(flight$levels)


str(flight)

```


#dropping the DEP_TIME column

```{r}

flight <- flight[, !colnames(flight) %in% c("DEP_TIME")]

```


#Train-Test Split 

```{r}

library(caret)

train_rows <- createDataPartition(y = flight$Flight.Status, p = 0.7,list = F)

train <- flight[train_rows,]

test <- flight[-train_rows,]

```


#Build the Naive Bayes Classifier

```{r}

library(e1071)

model_nb <- naiveBayes(train$Flight.Status~., train)

print(model_nb)

```


#Measure the model performance in train data

```{r}

preds <- predict(model_nb,train)

actuals <- train$Flight.Status

conf_mat_train = table(actuals,preds)

#Train metrics

Accuracy_tr = (conf_mat_train[1,1]+conf_mat_train[2,2])/sum(conf_mat_train)

Recall_tr = conf_mat_train[1,1]/(conf_mat_train[1,1]+conf_mat_train[2,2])

Specificity_tr = conf_mat_train[2,2]/(conf_mat_train[2,2]+conf_mat_train[2,1])

Precision_tr = conf_mat_train[1,1]/(conf_mat_train[1,1]+conf_mat_train[2,1])

f1_score_tr = (2*Precision*Recall)/(Precision+Recall)

output <- c(Accuracy_tr,Recall_tr,Specificity_tr,Precision_tr,f1_score_tr)

print(output)

```

#Accuracy/ Classification Rate - 81% - The percentage of flight arrivals and delays that are predicted correctly

#Recall - 96% - The flight delays correctly recognized

#Specificity - 95% - Flights that are on time are recognized correctly

#Precision - 82% - flights that are delayed are indeed delayed

#HIGH RECALL and LOW PRECISION - Flights that are delayed are predicted correctly. However, flights that are not delayed are predicted as delayed




```{r}
preds <- predict(model_nb, test)

actuals <- test$Flight.Status

conf_mat_test = table(actuals, preds)

#Train metrics

Accuracy_te = (conf_mat_test[1,1]+conf_mat_test[2,2])/sum(conf_mat_test)

Recall_te = conf_mat_test[1,1]/(conf_mat_test[1,1]+conf_mat_test[2,2])

Specificity_te = conf_mat_test[2,2]/(conf_mat_test[2,2]+conf_mat_test[2,1])

Precision_te = conf_mat_test[1,1]/(conf_mat_test[1,1]+conf_mat_test[2,1])

f1_score_te = (2*Precision*Recall)/(Precision+Recall)

output_test <- c(Accuracy_te,Recall_te,Specificity_te,Precision_te,f1_score_te)

print(output_test)

```

